---
title: "Milestone5"
author: "Mengxi Tan"
date: "10/18/2019"
output: html_document
---

**#About This Project**

Every year, there are roughly ~1M immigrants coming to the US from various origins and through different admission channels. How are these immigrants fitting in once they are in the US? Is there a certain a certain group that needs particular help in fitting into the American society?

One important metrics for fitting into the American society, is English speaking capability. According to the US census bureau, in 2017, while 91% of all US citizens self-identified as speaking English "very well", the number among all foreigh-born citizens was only 52%.

Would this impairment in English-speaking capability among immigrants differ by origin, channel of admission, or time? This visualization project aims to provide users with a tool to explore this issue, during the time span of 2007 - 2017.

*time span determined by data availability


**#Raw Data Source**

My raw data comes from two sources: 

1. Homeland Security: 
https://www.dhs.gov/immigration-statistics/yearbook

    covering annual immigration count by different characteristics (visa type, country of origin etc.)

2. American Fact Finder(US Census Bureau): https://factfinder.census.gov/faces/nav/jsf/pages/searchresults.xhtml?refresh=t#none

    covering US resident census data(immigration status, english speaking capability, in labor force or not, etc.)
    
Both data sets address the issue of immigrants to US, segmented by country and state of origin. Note that the US Census Bureau data covers all immigrants *living* in US in that particular year, while the Homeland Security data covers immigrants *moving* to US in that year. Therefore, there is strong connection between the two datasets, yet the Homeland Security sample population is only part of the US Census Bureau sample population.

**#Data Storage and Cleaning: for TF only, will not be shown to users**

Homeland Security data are stored under: raw_data > immigration_yearbook

US Census Bureau data are stored under: raw_data > english_speaking_immigrants

Relevant data that may or may not be used later is under: raw_data > parkinglot_4data

All data loading is finished at this stage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#When installing package data.table, choose "No compliation" method

library(readxl)
library(skimr)
library(ggplot2)
library(readxl)
library(janitor)
library(gifski)
library(gganimate)
library(data.table)
library(png)
library(plyr)
library(cowplot)
library(ggridges)
library(rstanarm)
library(broom)
library(tidyverse)

#####
#DATA LOADING: Homeland Security Part

# load the immigration by continent data
# clean up names and values
# filter out "Total" and "Unknown"

immigration_continent <- read_xlsx("raw_data/immigration_yearbook/Immigration_ByContinent.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  filter(region %in% c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")) %>%
  mutate(year = as.integer(year))

# rename columns
setnames(immigration_continent, "immediate_relatives_of_u_s_citizens", "immediate_relatives")

setnames(immigration_continent, "family_sponsored_preferences", "family")

setnames(immigration_continent, "employment_based_preferences", "employment")

setnames(immigration_continent, "refugees_and_asylees", "refugee")

# create a long format for continent level data
immigration_continent_long <- immigration_continent %>%
      select(region, year, immediate_relatives, family,employment, diversity, refugee, other) %>%
      gather(key = "admission_class", value = "count", immediate_relatives: other)

# load the immigration by country data
# clean up names and values

immigration_country_2017 <- read_xlsx("raw_data/immigration_yearbook/2017_table10d.xlsx", na = c("", "NA")) %>%
  clean_names() %>%
  mutate(year = 2017)

immigration_country_2016 <- read_xlsx("raw_data/immigration_yearbook/2016_table10d.xlsx", na = c("", "NA")) %>%
  clean_names() %>%
  mutate(year = 2016)

immigration_country_2015 <- read_xlsx("raw_data/immigration_yearbook/2015_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2015)

immigration_country_2014 <- read_xlsx("raw_data/immigration_yearbook/2014_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2014)

immigration_country_2013 <- read_xlsx("raw_data/immigration_yearbook/2013_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2013)


immigration_country_2012 <- read_xlsx("raw_data/immigration_yearbook/2012_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2012)

immigration_country_2011 <- read_xlsx("raw_data/immigration_yearbook/2011_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2011)

immigration_country_2010 <- read_xlsx("raw_data/immigration_yearbook/2010_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2010)

immigration_country_2009 <- read_xlsx("raw_data/immigration_yearbook/2009_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2009)

immigration_country_2008 <- read_xlsx("raw_data/immigration_yearbook/2009_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2008)

immigration_country_2007 <- read_xlsx("raw_data/immigration_yearbook/2009_table10d.xlsx", na = c("", "NA")) %>%
  clean_names()%>%
  mutate(year = 2007)


#Combine the country tables together
immigration_country <- rbind(immigration_country_2007,
                             immigration_country_2008,
                             immigration_country_2009,
                             immigration_country_2010,
                             immigration_country_2011,
                             immigration_country_2012,
                             immigration_country_2013,
                             immigration_country_2014,
                             immigration_country_2015,
                             immigration_country_2016,
                             immigration_country_2017)

# clean up table
# convert year to integer for graphing later
immigration_country <- immigration_country %>%
                          filter(region_and_country_of_birth != "Total") %>%
                          drop_na %>%
                          mutate(year = as.integer(year))

# rename columns
setnames(immigration_country, "region_and_country_of_birth", "country")

setnames(immigration_country, "immediate_relatives_of_u_s_citizens", "immediate_relatives")

setnames(immigration_country, "family_sponsored_preferences", "family")

setnames(immigration_country, "employment_based_preferences", "employment")

setnames(immigration_country, "refugees_and_asylees", "refugee")

# "-" means 0, D means cannot disclose
# treat both as 0 in this case
# convert to numeric value

# family column
immigration_country$family <- revalue(immigration_country$family, c("-"= "0"))
immigration_country$family <- revalue(immigration_country$family, c("D"= "0"))
immigration_country$family <- as.numeric(immigration_country$family)

# employment column
immigration_country$employment <- revalue(immigration_country$employment, c("-"= "0"))
immigration_country$employment <- revalue(immigration_country$employment, c("D"= "0"))
immigration_country$employment <- as.numeric(immigration_country$employment)

# relative column
immigration_country$immediate_relatives <- revalue(immigration_country$immediate_relatives, c("-"= "0"))
immigration_country$immediate_relatives <- revalue(immigration_country$immediate_relatives, c("D"= "0"))
immigration_country$immediate_relatives <- as.numeric(immigration_country$immediate_relatives)

# diversity column
immigration_country$diversity <- revalue(immigration_country$diversity, c("-"= "0"))
immigration_country$diversity <- revalue(immigration_country$diversity, c("D"= "0"))
immigration_country$diversity <- as.numeric(immigration_country$diversity)

# refugee column
immigration_country$refugee <- revalue(immigration_country$refugee, c("-"= "0"))
immigration_country$refugee <- revalue(immigration_country$refugee, c("D"= "0"))
immigration_country$refugee <- as.numeric(immigration_country$refugee)

# other column
immigration_country$other <- revalue(immigration_country$other, c("-"= "0"))
immigration_country$other <- revalue(immigration_country$other, c("D"= "0"))
immigration_country$other <- as.numeric(immigration_country$other)

# create a long format for continent level data
immigration_country_long <- immigration_country %>%
      select(country, year, immediate_relatives, family,employment, diversity, refugee, other) %>%
      gather(key = "admission_class", value = "count", immediate_relatives: other)

# rearrange the order of categorical variables for labelling later
immigration_country_long$admission_class <- factor(immigration_country_long$admission_class, levels = c("diversity", "employment", "immediate_relatives", "family", "refugee", "other"))

# make a clean data directory
dir.create("clean_data")

######
#DATA LOADING: US Census Bureau Part
#load english speaking capability by continent

# english speaking: only takes into consideration population 5+
# employment status: only takes into consideration population 16+

english_by_continent  <- read_xlsx("raw_data/social_immigrants/english_by_continent.xlsx", na = c("", "NA")) %>% clean_names()

#load english speaking capability by country, for the top 10 country with most US immigrants
english_by_country <-
read_xlsx("raw_data/social_immigrants/english_by_country.xlsx", na = c("", "NA")) %>% clean_names()

# load unemployment by country, for the top 10 country with most US immigrants
employment_by_country <-
read_xlsx("raw_data/social_immigrants/employment_by_country.xlsx", na = c("", "NA")) %>% clean_names()

# write the clean data to csv
write.csv(immigration_country, file = "clean_data/immigration_country.csv", row.names = FALSE)
write.csv(immigration_continent, file = "clean_data/immigration_continent.csv", row.names = FALSE)
write.csv(immigration_country_long, file = "clean_data/immigration_country_long.csv", row.names = FALSE)
write.csv(immigration_continent_long, file = "clean_data/immigration_continent_long.csv", row.names = FALSE)
write.csv(english_by_continent, file = "clean_data/english_continent.csv", row.names = FALSE)
write.csv(english_by_country, file = "clean_data/english_country.csv", row.names = FALSE)
write.csv(employment_by_country, file = "clean_data/employment_country.csv", row.names = FALSE)

```


```{r investigation/graph making - testing for app.R}

# examine the top 10 countries with the most immigrants
# Mexico 
# China, People's Republic
# India 
# Philippines 
# Dominican Republic
# Cuba 
# Vietnam  
# Korea, South
# Colombia
# Haiti
top10 <- immigration_country %>%
          group_by(country) %>%
          summarize(total = sum(total)) %>%
          arrange(desc(total))
          head(10)

# create a intro_graph

intro_graph <- ggplot(immigration_country_top_10, aes(x = total, y = country, group = country)) + geom_density_ridges(fill = "orange") + labs(x = "Total Immigrants", y = "Country", title = "Top 10 Countries with Most Immigration into US ", caption = "2009 - 2017")

# save for later usage
ggsave("Immigration_Explorer/intro_graph.png", intro_graph)



  



  
  


```

